---
title: "DATA 607 - WEEK 2"
author: "Paul Britton"
output:  
  html_document:
    theme: cerulean
    highlight: pygments

---

My friends and family have probably had enough of me talking about data, so rather than interview them, I decided to use a movie dataset that I found on the internet [here](https://grouplens.org/datasets/movielens/).  I've copied the data to my [github](https://github.com/plb2018/DATA607/tree/master/ml-latest-small) and reference that copy in my code.  

I've opted to do all my MySQL work directly from R, so the only file that i have to show is a .rmd which can be found in my [here](https://github.com/plb2018/DATA607/blob/master/DATA607_week2.Rmd) in my github and [here on rpubs](http://rpubs.com/plb_lttfer/359133)


### Load Required Packages

```{r results='hide', message=FALSE, warning=FALSE}
library(RMySQL)
library(data.table)

```


### Read in username and password from file

```{r results='hide'}

user_df <- read.table("mysql_temp_pw.csv",header=TRUE,sep=",")

user = toString(user_df[1,"user"])
pw = toString(user_df[1,"password"])

```

First we'll open a connection to MySQL. Note that I've pre-created a database called "DATA_607" using the following SQL command: CREATE DATABASE "DATA_607".

```{r}
con = dbConnect(MySQL(), user=user, password=pw, dbname='DATA_607', host='localhost')


```


Next we'll read the movie data into dataframes.  The data we care about are contained in 2 separate files (ratings, movies) which we'll end up combining into a single db table later.  

```{r}


ratings <- read.table("https://raw.githubusercontent.com/plb2018/DATA607/master/ml-latest-small/ratings.csv",header = TRUE, sep = ",")

movies <- read.table("https://raw.githubusercontent.com/plb2018/DATA607/master/ml-latest-small/movies.csv",header = TRUE, sep = ",",fill = TRUE)

```

And we'll take a look at the data:

```{r}

head(ratings,5)

```

```{r}

head(movies,5)

```


The data look good, so now we'll load the data into individual tables in MySQL:


```{r results='hide'}
dbWriteTable(con, name='ratings', value=ratings)
dbWriteTable(con, name='movies', value=movies)

```

Now we should have 2 tables: "movies" and "ratings".  Let's check:

```{r}

dbListTables(con)
```

Looks good!  Now we'll pull some data and take a peek just to make sure that everything is as we expect:

```{r}

query = dbSendQuery(con, "SELECT * FROM movies LIMIT 5")
movie_data = fetch(query)
movie_data
```

```{r}

query = dbSendQuery(con, "SELECT * FROM ratings LIMIT 5")
rating_data = fetch(query)
rating_data
```

The tables look fine, so we'll try to join the data "movies" and "ratings" tables using the common "movieId" field and create a new "movie_ratings" table to house the joined data:



```{r results='hide'}

dbSendQuery(con, "CREATE TABLE movie_ratings AS (
SELECT  movies.movieId,ratings.userId, ratings.rating, movies.title,movies.genres 
FROM movies
INNER JOIN ratings ON  movies.movieId = ratings.movieId);")

```
And finally, we'll pull a sample out of the newly created table into an R dataframe:

```{r}

query = dbSendQuery(con, "SELECT * FROM movie_ratings LIMIT 5")
data = fetch(query)

data
```

And finally, we'll close out db conneciton.


```{r results='hide', message=FALSE, warning=FALSE}
dbDisconnect(con)

```
