---
title: "DATA 607 - WEEK 9 Assignment"
author: "Paul Britton"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
  pdf_document:
    toc: yes
---

```{r, echo=FALSE}
rm(list = ls())
```


The task for week 9 was to get a NYT API key, construct an interface to R, grab some data and finally return a dataframe.  I chose to work with the "Archive_api" because I actually have some future tasks that I'd like to tackle that involve old newspapers.

This file is available on rpubs [here](http://rpubs.com/plb_lttfer/375108) and in my github [here]()


### Load the Libraries

```{r, message=FALSE, warning=FALSE}

library(jsonlite)
library(httr)
library(knitr)

```


### Build a function to Get JSON from NYT Archives API

I created a function to query the NYT Archive API and return a dataframe containing the data retrieved.  Also, in the development process, I noticed that my request failed frequently so I've set my function up to retry using the RETRY() function.  

```{r}

#params
api.key <- "36a5b43cb0e04a1dad5e23a9810f2cc1"
yyyy <- "1929"
mm <- "09"

#return JSON from NYT API
get.NytArchives <- function(api.key,yyyy,mm){
  base.url <-  paste("https://api.nytimes.com/svc/archive/v1/",yyyy,"/",mm,".json",sep="")
  print(paste("Collecting NYT archvies data for: ",toString(yyyy),"-",toString(mm)))
  
  #get seems to fail sometimes, so keep on tryin'
  query <- RETRY("GET","https://api.nytimes.com/svc/archive/v1/1929/9.json",
                 query = list(api_key=api.key),
                 times = 100, 
                 pause_base = 2)
  query <- content(query,as="text",encoding="UTF-8")
  
  df <-  as.data.frame(fromJSON(query))
  
  #clean up the column names
  colnames(df) <- gsub("^.*\\.","", colnames(df))
  
  return(df)
}

```

### Test the Function

We'll do a call to grab a single month and see what we get back

```{r}

result <- get.NytArchives(api.key,"1929","9")

```



### Check the Output

And now we'll take a look and see what we got.  First i'll print the column names:  

```{r}

kable(colnames(result),col.names = "Column Names")
```

For output purposes, I'll select only a few of the columns listed above in order to keep things legible:

```{r}
kable(head(result[c("web_url","snippet")],5))

```


The data looks good!


Now we're going to try and grab a bunch of data all at once:

```{r}

df <- data.frame(matrix(ncol = 2, nrow=0))

colnames(df) <- c("web_url","snippet")

for (i in 1:5){
  data <- get.NytArchives(api.key,1929,i)
  
  df<- rbind(df,data[c("web_url","snippet")])
}




```


We've just collected `r nrow(df)` articles from the NYT archives

