---
title: "Data 607 - Week 1"
author: "Paul Britton"
date: '2018-02-04'
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

The .rmd and raw data for this assignment can be found in my github repository located at [https://github.com/plb2018/DATA607](https://github.com/plb2018/DATA607). For the purposes of this exercise, i'm grabbing the data files directly from my github.

The rpubs.com version of this work can be found at [http://rpubs.com/plb_lttfer/](https://github.com/plb2018/DATA607)


The task here is to perform several manipulations of a mushroom data set so first, we'll get the data, grab the first few columns and take a look:

###Load data

```{r load_data, results='hide'}
library(data.table)
library(plyr)

#load the data
shrooms <- fread("https://github.com/plb2018/DATA607/raw/master/agaricus-lepiota.data")

```

###Downscope the data and inspect

```{r downscope_data}

#grab the first few columns 
df <- data.frame(shrooms[,0:6])

head(df,5)

```



###Name the columns

The data looks good, so now we give the columns some names:

```{r name_cols}
#rename the columns
names(df) <- c("edible_poisonous","cap_shape","cap_surface","cap_color","bruises","odor")

head(df,5)

```


###Replace abbreviations

The re-naming of the columns looks good, so now we'll replace the abbreviations with the appropriate names.  The source file for the names can be found [here](https://github.com/plb2018/DATA607/raw/master/agaricus-lepiota.names):

```{r abbrv_to_names}

df$edible_poisonous <- mapvalues(df$edible_poisonous,
                                 from=c('e','p'),
                                 to=c('edible','poisonous'))



df$cap_shape <- mapvalues(df$cap_shape,
                          from=c('b','c','x','f','k','s'),
                          to=c('bell','conical','convex','flat','knobbed','sunken'))



df$cap_surface <- mapvalues(df$cap_surface,
                            from=c('f','g','y','s'),
                            to=c('fibrous','grooves','scaly','smooth'))



df$cap_color <- mapvalues(df$cap_color,
                          from=c('n','b','c','g','r','p','u','e','w','y'),
                          to=c('brown','buff','cinnamon','gray','green','pink','purple','red','white','yellow'))



df$bruises <- mapvalues(df$bruises,
                        from=c('t','f'), 
                        to=c('bruises','no bruises'))



df$odor <- mapvalues(df$odor,
                     from=c('a','l','c','y','f','m','n','p','s'),
                     to=c('almond','anise','creosote','fishy','foul','musty','none','pungent','spicy'))



head(df,5)

```

We now have the first few columns of the dataframe with meaningful names, and all abbreviations replaced!




