---
title: "DATA 607 - Project 1"
author: "Paul Britton"
output: 
  html_document:
    toc: TRUE

---


### Load Libraries

Load a few libraries that we will use along the way

```{r} 
library(stringr) 
library(tidyr)
library(ggplot2)
``` 


### Load the Data

First we load the data ant take a look

```{r} 

source_data<- "https://raw.githubusercontent.com/plb2018/DATA607/master/DATA_607_Project1/tournamentinfo.txt" 
rankings <- unlist(read.table(source_data,header=FALSE,sep="\n")) 
head(rankings,10)
```

We can see immediately that the data needs a bit of processing before we can do anything with it.  We'll try to parse it with some regular expressions

### Parse the Data

It looks like we can extract what we need using regular expressions.  The data that we need to grab is spread across 2 rows for each player.  In both cases, the fields are consistently separated by pipes, making things a bit easier.  The first row (r1) that we want to grab begins with the player number followed by a space and a "|" delimiter.  The second row (r2) that we want begins with the the state-code (2 uppercase letters), followed by a space and a "|" delimiter.  This information should  be sufficent for us to grab and recombine the rows in a more usable format.   I sometimes find regular expressions hard to decipher, particularly if i haven't worked with them in a while, so i tend to explicitly explain them in my comments.


```{r} 


#if we see a 1-4 digit # followed by a space and a pipe, grab the it + the rest of the row
r1 <- unlist(str_extract_all(rankings[5:length(rankings)],"\\d{1,4}\\s\\|.+")) 

#if we see 2 uppercase letters followed by a space and a pipe, grab it + the rest of the row
r2 <- unlist(str_extract_all(rankings[5:length(rankings)],"[[:upper:]]{2}\\s\\|.+")) 

#while we're at it, we'll replace the "W's" "L's" "D's" and "B's" in r1... we just want thte #s
r1 <- unlist(str_replace_all(r1,"\\|(W|L|B|D|H|U)\\s","|"))


head(r1,5)
head(r2,5)


```

The regex appears to have works as intended, so now we'll split the rows on the "|"'s while loading the data to a dataframe.  We'll then  down-select to just the cols that we need.

```{r} 

#create the df
df <- data.frame(str_split_fixed(r1,"\\|",n=11),str_split_fixed(r2,"\\|",n=10))


#grab the data we need
cols <- c(1:10,12:13)
df <- df[,cols]

#re-order the cols 
cols <- c(1,2,11,12,3:10)
df <- df[,cols]



#we'll also rename all the cols that we're going to keep while we're at it
names(df) <- c("PlayerNum","PlayerName","PlayerState",
               "USCF_Pre_Post","TotalPts","r1","r2",
               "r3","r4","r5","r6","r7")

#trim the whitespace
df <- as.data.frame(apply(df,2,function(x)gsub("^\\s+|\\s+$", "", x)))

head(df,5)
``` 


Now the data looks reasonably close to where we want it to be, however, we still need to deal with that last column (ID_Pre_Post).  Because the columns appear to be well formated, We're going to split the column into 3 using separate() from tidyr.  We're really only after the pre_rating column here.  Another thing of note is that the pre-rating column sometimes contains a "p" with the ranking that we want consistently on the lefthand side of the "p".

```{r warning=FALSE} 

#split the columns 
df <- separate(data=df,col=USCF_Pre_Post,sep=" / R: ",into = c("USCF ID","Pre_Post"))
df <- separate(data=df,col=Pre_Post,sep="   ->",into = c("Pre_Rating","Post_Rating"))

#pre_rating sometimes contains a "p".  the resultant "etc" column is a throw-away
df <- separate(data=df,col=Pre_Rating,sep="P",into = c("Player_Pre_Rating","etc"))

head(df,5)

``` 

### Compute the Pre-Chess Rating


The table looks good - all i have to do now is compute the average pre-chess score and drop all the unwanted columns.  To compute the average score, i'm going to loop over each player, figure out who they played against (cols 9:15) and use that data to look-up the pre-game ratings.  Using that data, i can compute the averages.


```{r warning=FALSE} 

pre_chess_rating = c()

for (i in 1:nrow(df)){

  #get all players who played against this player
  opponent_nums <- df[i,9:15]
  opponent_nums <-  as.numeric(levels(unlist(opponent_nums)))[unlist(opponent_nums)]
  opponent_nums <- opponent_nums[!is.na(opponent_nums)]
  
  #pull scores of the opponents and compute the avg
  pre_chess_rating[i] <- trunc(mean(as.numeric(df[opponent_nums,5])))  
 
}

df$Player_Pre_Rating <- as.numeric(df$Player_Pre_Rating )

df$Avg_Opponent_Rating = pre_chess_rating


```

### Write the File

Now we drop all the unwanted cols and write the data to file:

```{r warning=FALSE} 

df <- df[c(2:3,8,5,16)]

head(df)

write.csv(df, file = "data_607_project1.csv")

``` 


### Bonus Plots

I just wanted to take a look and see whether there were any obvious relatioships in the data.  We see below that there appears to be a positive relatioship between total points and player_pre_rating.  I also wanted to  see if i could extract any info by adding an additional dimension to the plot (avg_opponent_rating) via dot-color.    

```{r warning=FALSE} 

ggplot(df, aes(Player_Pre_Rating,TotalPts)) + geom_point(aes(color=Avg_Opponent_Rating)) +
ggtitle("")


``` 